Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-08-26 doc/papers/cgo2014/paper.tex
Unifiy StackOverflow (camel) case

    This papers focuses on the case of the Python language and especially the
    Numpy package that provides core array data structure and basic linear
    algebra routines. It first conducts a case study based on Numpy use
    cases from the stackoverflow question and answer website, then studies
    existing compilers for numerical Python. Optimization opportunities,
    including parallelization and vectorization, and their implementation in
    the Pythran compiler are presented and illustrated through several
    benchmarks, showing very interesting speedups over the standard Python
    interpreter. Speedup greater than a factor of 10 are achieved despite the
    fact that the considered benchmarks mostly call Numpy routines that run C code.

@@

This paper focuses on compilation techniques that are relevant for the
optimization of high-level numerical kernels written in Python using the Numpy
package. Section~\ref{sec:stackoverflow} presents a case study based on user
input gathered on the stackoverflow question and answer site.
Section~\ref{sec:optimize} focuses on a simple yet representative kernel and
showcases the optimization opportunities typically found in Numpy-based
kernels. Existing compilation approach to optimize such kernels are discussed
in Section~\ref{sec:compilers}, then Section~\ref{sec:pythran} emphasises on a
compiler-runtime cooperative approach to solve the performance issues.
Experimentations showcasing important performance gains are detailed in
Section~\ref{sec:xp}.

@@

One of the first steps required when designing an optimizing compiler is to
gather enough test cases, or benchmarks, in order to drive the optimization
process toward realistic examples. To achieve this goal, we have gathered a few
example from existing compilers' testsuite, but these may have been modified to
overcome some compiler limitations. A less biased source of benchmarks should
be independant from the compiler community and focused on matching user needs.
To achieve that goal, we collected synthetic scientifc kernels from the
stackoverflow website.

>>>>

    This papers focuses on the case of the Python language and especially the
    Numpy package that provides core array data structure and basic linear
    algebra routines. It first conducts a case study based on Numpy use
    cases from the StackOverflow question and answer website, then studies
    existing compilers for numerical Python. Optimization opportunities,
    including parallelization and vectorization, and their implementation in
    the Pythran compiler are presented and illustrated through several
    benchmarks, showing very interesting speedups over the standard Python
    interpreter. Speedup greater than a factor of 10 are achieved despite the
    fact that the considered benchmarks mostly call Numpy routines that run C code.

@@

This paper focuses on compilation techniques that are relevant for the
optimization of high-level numerical kernels written in Python using the Numpy
package. Section~\ref{sec:stackoverflow} presents a case study based on user
input gathered on the StackOverflow question and answer site.
Section~\ref{sec:optimize} focuses on a simple yet representative kernel and
showcases the optimization opportunities typically found in Numpy-based
kernels. Existing compilation approach to optimize such kernels are discussed
in Section~\ref{sec:compilers}, then Section~\ref{sec:pythran} emphasises on a
compiler-runtime cooperative approach to solve the performance issues.
Experimentations showcasing important performance gains are detailed in
Section~\ref{sec:xp}.

@@

One of the first steps required when designing an optimizing compiler is to
gather enough test cases, or benchmarks, in order to drive the optimization
process toward realistic examples. To achieve this goal, we have gathered a few
example from existing compilers' testsuite, but these may have been modified to
overcome some compiler limitations. A less biased source of benchmarks should
be independant from the compiler community and focused on matching user needs.
To achieve that goal, we collected synthetic scientifc kernels from the
StackOverflow website.
